They are Very Smart: Genetic-Algorithm-Based Pathfinding in Games

### 2018-12-11 paper.tex
More work on paper

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned
onslaught of enemies such as the fearsome Zombunny (see Figure 3). Zombunnies
are driven by a single goal: destruction of the player's home base. Zombunnies
spawn in groups of ten (separated by a slight delay) at the beginning of the
game. Subsequent waves spawn after all members from the previous wave have died.
Zombunnies use a genetic algorithm in conjunction with a recurrent neural
network to evolve a pathfinding algorithm between generations (waves), with the
long-term incentive of dealing damage to the player's base% (explain fitness
% criteria more concisely here)
. Zombunnies can be killed by either taking damage
from a player's defenses or through "starvation." A Zombunny will starve to
death if it does not take or deal damage for a given length of time. Through
this starvation mechanism, the Zombunnies which survive the longest are
ultimately those which interact the most with the player's defenses, and are
subsequently more likely to pass on attributes of their genome to the next
generation. Zombunnies have three characteristic attributes: health, damage, and
movement speed. All three of these stats are assigned upon spawning as dictated
by their genome. Pathfinding is controlled by a recurrent neural net, whose
weights are determined through a genetic algorithm between waves, which
optimizes for Zombunnies' ability to achieve their overarching goal of player
base destruction. Zombunnies can "see" nearby entities within a certain radius,
including other Zombunnies, player defenses, and the player's home base (when in
range). Based on the combination of visible inputs, Zombunnies will output a
chosen direction which they will then walk in. This algorithm updates on a
tick-by-tick basis. Zombunnies will damage structures they collide with. Because
evolution of the pathfinding neural net only happens between generations, near
the beginning of the game, Zombunnies will start off with relatively stupid
pathfinding which will typically cause the first few populations to starve to
death without dealing any significant damage. This provides the player with an
initial grace period during which they may build their initial defenses. As time
goes on, Zombunnies will evolve towards an optimal stat distribution of health,
damage, and speed, as well as in their ability to pathfind in such a way as to
thwart the player's defenses and end the game.

@@

Insert picture of \textit{They Are Billions} gameplay here.

@@

The Spawn method, as seen in Figure 12, is called every SPAWN\_DELAY interval POPULATION\_COUNT times. The function starts by spawning a Zombunny with no genome due to the inability to provide parameters upon GameObject instantiation. Therefore, we set the genome after the fact, popping it off of the toSpawn stack, also adding the ZombieDied function to the removed event on said Zombunny, which allows us to execute code in the context of the GenomeManager upon Zombunny death; namely, ZombunnyDied. ZombunnyDied handles finding the fitness of Zombunnies and adding them to the list of completed genomes, as well as calling NextWave again when the last Zombunny in the wave dies.

@@

Redzen is a general-purpose utilities library for C\# that contains the fast random function that we use. However, it should be noted that the library as it may be found online at the time of writing does not compile with the older C\# compiler that Unity uses. Thus, we had to manually delete most of the library files and remove underscore digit separators from some numeric constants, as they are a C\# 7 feature. You can find our compatible library at \path{They-Are-Very-Smart/TD/TowerDefense/Assets/Scripts/Util/Redzen}

@@

ZombieAgent is a GameObject which determines Zombunny behavior and stats from a genome and keeps track of fitness for each individual Zombunny. To do so, ZombieAgent pulls all the stat-determining values from the genome and applies them upon instantiation, i.e. setting health, damage, and speed as determined by the genome, as well as handling tick-to-tick operations during a Zombunny's lifetime. These tick-to-tick operations include choosing the next movement vector based on the location of nearby GameObjects and agent's current status.

\imfig{ZombieAgent1}{ZombieAgent class parameters}

@@

\imfig{ZombieAgentVision}{The methods for keeping track of GameObjects entering
  and exiting a Zombunny's visionCollider}

@@

\imfig{ZombieAgentUpdateNearby}{The portions of the FixedUpdate method for
determining the position of nearby objects relative to a Zombunny}

@@

Figure 17 shows the portion of FixedUpdate dedicated to rotating a Zombunny from
tick to tick, calling the Rotate method of the lastVelocity PolarVector using
the CalculateDirection method of the genome class, taking into account the
Zombunny's previous velocity vector, the time alive, the Zombunny's current
health, and the aforementioned dictionary of nearby GameObjects.

@@

>>>>

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned
onslaught of enemies such as the fearsome Zombunny (see Figure 4). Zombunnies
are driven by a single goal: destruction of the player's home base. Zombunnies
spawn in groups of ten (separated by a slight delay) at the beginning of the
game. Subsequent waves spawn after all members from the previous wave have died.
Zombunnies use a genetic algorithm in conjunction with a recurrent neural
network to evolve a pathfinding algorithm between generations (waves), with the
long-term incentive of dealing damage to the player's base% (explain fitness
% criteria more concisely here)
. Zombunnies can be killed by either taking damage
from a player's defenses or through "starvation." A Zombunny will starve to
death if it does not take or deal damage for a given length of time. Through
this starvation mechanism, the Zombunnies which survive the longest are
ultimately those which interact the most with the player's defenses, and are
subsequently more likely to pass on attributes of their genome to the next
generation. Zombunnies have three characteristic attributes: health, damage, and
movement speed. All three of these stats are assigned upon spawning as dictated
by their genome. Pathfinding is controlled by a recurrent neural net, whose
weights are determined through a genetic algorithm between waves, which
optimizes for Zombunnies' ability to achieve their overarching goal of player
base destruction. Zombunnies can "see" nearby entities within a certain radius,
including other Zombunnies, player defenses, and the player's home base (when in
range). Based on the combination of visible inputs, Zombunnies will output a
chosen direction which they will then walk in. This algorithm updates on a
tick-by-tick basis. Zombunnies will damage structures they collide with. Because
evolution of the pathfinding neural net only happens between generations, near
the beginning of the game, Zombunnies will start off with relatively stupid
pathfinding which will typically cause the first few populations to starve to
death without dealing any significant damage. This provides the player with an
initial grace period during which they may build their initial defenses. As time
goes on, Zombunnies will evolve towards an optimal stat distribution of health,
damage, and speed, as well as in their ability to pathfind in such a way as to
thwart the player's defenses and end the game.

@@

%Insert picture of \textit{They Are Billions} gameplay here.%

@@

The Spawn method, as seen in Figure 12, is called at every SPAWN\_DELAY interval POPULATION\_COUNT times. The function starts by spawning a Zombunny with no genome due to the inability to provide parameters upon GameObject instantiation. Therefore, we set the genome after the fact, popping it off of the toSpawn stack, also adding the ZombieDied function to the removed event on said Zombunny, which allows us to execute code in the context of the GenomeManager upon Zombunny death; namely, ZombunnyDied. ZombunnyDied handles finding the fitness of Zombunnies and adding them to the list of completed genomes, as well as calling NextWave again when the last Zombunny in the wave dies.

@@

Redzen is a general-purpose utilities library for C\# that contains the fast random function that we use. However, it should be noted that the library as it may be found online at the time of writing does not compile with the older C\# compiler that Unity uses. Thus, we had to manually delete most of the library files and remove underscore digit separators from some numeric constants, as they are a C\# 7 feature. You can find our compatible library at \path{They-Are-Very-Smart/TD/TowerDefense/Assets/Scripts/Util/Redzen}.

@@

ZombieAgent is a GameObject which determines Zombunny behavior and stats from a genome and keeps track of fitness for each individual Zombunny. To do so, ZombieAgent pulls all the stat-determining values from the genome and applies them upon instantiation, i.e. setting health, damage, and speed as determined by the genome, as well as handling tick-to-tick operations during a Zombunny's lifetime. These tick-to-tick operations include choosing the next movement vector based on the location of nearby GameObjects and agents' current status.

\imfigH{ZombieAgent1}{ZombieAgent class parameters}

@@

\imfigH{ZombieAgentVision}{Methods for keeping track of GameObjects entering
  and exiting a Zombunny's visionCollider}

@@

\imfigB{ZombieAgentUpdateNearby}{The portion of the FixedUpdate method for
determining the position of nearby objects relative to a Zombunny}

@@

Figure 17 shows the portion of FixedUpdate dedicated to altering a Zombunny's
trajectory from tick to tick, calling the Rotate method of the lastVelocity
PolarVector using the CalculateDirection method of the genome class, taking into
account the Zombunny's previous velocity vector, time alive, current health, and
the aforementioned dictionary of nearby GameObjects.

@@

\subsection{Evolution}

In \
